HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
edits

This section summarises the manual annotation procedure performed in order 
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.

@@

In non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Since errors in morphology
should be expressed on the atomic units, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish
compared to German and Romanian\oa{but doesn't German also have a case system? Ondrej:
  could you pinpoint what the difference exactly is?}.

An example HUME annotation is given in \figref{fig:hume_tree},
which presents an annotation of an UCCA tree, and a corresponding translation.
We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
individually.

@@

\paragraph{Monolingual vs. Bilingual Evaluation}
HUME diverges from HMEANT (and also from the shallow automatic measures
like BLEU or METEOR) in not requiring a reference.
Instead, it compares the source directly with the output translation.

>>>>

This section summarises the manual annotation procedure performed in order 
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

@@

In non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Errors in morphology
should be expressed on the atomic units, so HUME can behave slightly differently
on configurational vs. non-configurational languages.

% An example HUME annotation is given in \figref{fig:hume_tree},
% which presents an annotation of an UCCA tree, and a corresponding translation.
% We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
% in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
% individually.

@@

\paragraph{Monolingual vs. Bilingual Evaluation}
\label{src-vs-ref}
HUME diverges from HMEANT (and also from the shallow automatic measures
like BLEU or METEOR) in not requiring a reference.
Instead, it compares the source directly with the output translation.
